---
title: "Data Camp Intro to R Assignment"
author: "Aliza Fassler"
date: "9/11/2021"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Q1
(1pt): What type of data is contained in the variable a
```{r}
a<- "Aliza"
class(a)
```
Variable a is a character.

# Q2
(1pt): What type of data is contained in the variable b1

```{r}
b1<-45.6
b1
class(b1)
```
Variable b1 is numeric.

# Q3
(1pt): What type of data is contained in the variable b2
```{r}
b2<-"45.6"
```
Variable b2 is a character.

# Q4
(2pt): What happens when you add b1 and b2 and why?

When you add b1 + b2 you get an error message because you cannot add variables of different data types. B1 is a number (numeric) and b2 is a character (text) variable.

# Q5
(1pt): Are the variables b1 and c1 of the same type?
```{r}
c1<-c(0:3)
c1
class(c1)
```
Variables b1 and c1 are not the same data type. B1 is numeric meaning it can have decimal values. C1 is an integer meaning it can only have whole numbers.

# Q6
(3pt): Explain what happens when you add b1 and c1. Consider both the number of elements in each variable and the data types.
```{r}
b1+c1
```

When you add b1 + c1 because there is only 1 value in b1 it adds b1 to each value in c1.It treats all the values in b1 and c1 as numeric in this case. 

# Q7
(1pt): Show the R code you used to create v1.
```{r}
v1<-c(-2:2)
v1
```
# Q8
(1pt): Show the R code you used to create v2.
```{r}
v2<-v1*3
v2
```
# Q9
(1pt): Show the R code you used to calculate the sum of elements in v2.
```{r}
sum(v2)
```
# Q10
(1 pt) Show the code you used to create mat_1
```{r}
vec_4<-c(1:12)
vec_4
mat_1<-matrix(vec_4, byrow = TRUE, nrow = 3, ncol = 4)
mat_1
```
# Q11
(1 pt) Show the code you used to create mat_2

```{r}
mat_2<-matrix(vec_4, byrow = FALSE, nrow = 3, ncol = 4)
mat_2
```

# Q12
Q1 (2 pts): Show the R code you used to create my_list_1.
```{r}
list_vec<-0:5
my_list_1<-list(5.2, "five point two", list_vec )
my_list_1
names(my_list_1)<-c("two", "one", "three")
```
# Q13
(1pt): Show the R code that would select third element of the list.
```{r}
my_list_1[[3]]
```
# Q14
Q2 (1pt): Show the R code that selects the list element with the name “one”. Note: there are at least two ways to do this!
```{r}
my_list_1$one
my_list_1[["one"]]
```

# Q15
(3 pts.) Show the R code that you used to create my_bool_vec.
```{r}
my_vec = rep(1:3, 5)
my_vec
```

# Q16
(2 pts.) Show the R code that you used to subset my_vec using my_boolean_vec.
```{r}
my_bool_vec<-my_vec == 3
data.frame(my_vec, my_bool_vec)
```

