---
title: "Lab 6"
author: "Aliza Fassler"
date: "10/13/2021"
output: html_document
---

I worked with Sonja Glasser and Justin Roch.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
require(here)
require(dplyr)
dat_pen = require(palmerpenguins)
```


# **Q1 (3 pts.): Show the R code you used to define your sse_mean() function. Include the following line before your function definition:**

```{r}
rm(list = ls())
sse_mean = function(x)
 { 
  y = na.omit(x)
  return(sd(y)/sqrt(length(y)))
}

sse_mean(penguins$bill_depth_mm)
sse_mean(mtcars$mpg)
```
#This is also an option
```{r}
rm(list = ls())
sse_mean2 = function(x)
 { 
  y = x[is.na(x) != 1]
  return(sd(y)/sqrt(length(y)))
}

sse_mean2(penguins$bill_depth_mm)
sse_mean2(mtcars$mpg)
```
```{r, include = FALSE}
dat_pen = droplevels(subset(penguins, species != "Gentoo"))
```


# **Q2 (4 pts.): Include the code you used to define your two_group_resample() function into the text input.**
```{r}
x = dat_pen$flipper_length_mm
n_1 = 68
n_2 = 152

two_group_resample = function(x, n_1, n_2) 
{
dat_1 = sample(x, n_1, replace = TRUE)
dat_2 = sample(x, n_2, replace = TRUE)

difference_in_means = mean(dat_1, na.rm = TRUE) - mean(dat_2, na.rm = TRUE)
  
  return(difference_in_means)
}
```

##testing function
```{r}
set.seed(54321)
two_group_resample(dat_pen$flipper_length_mm, 68, 152)
```
# **Q3 (2 pts.): Does your function perform Monte Carlo or bootstrap resampling, i.e. does it simulate a null or an alternative hypothesis? You may want to review your answer after you complete the rest of the lab questions.**

My function performs Monte Carlo function, because it resamples flipper length with no relationship to the species. This simulates the null hypothesis. 

# **Q4 (2 pts.): Create histogram of the resampled differences of means**

```{r}
n = 2000
mean_differences = c()
for (i in 1:n)
{
  mean_differences = c(
    mean_differences,
    two_group_resample(dat_pen$flipper_length_mm, 68, 152)
  )
}
hist(mean_differences, main = "Histogram of Simulated Mean Differences (Monte Carlo)", xlab = "Mean Differences")
```

# **Q5 (6 pts.): How many of your resampled differences of means were greater than 5.8?**

None because my sampling method approximated the null hypothesis. 

**Assume that we are doing a 2-sided test: we don’t care which species has longer** **flippers, we only want to know if they are different.**
**Include the R code you used to check the number of differences greater than 5.8.**

```{r}
t_test = t.test(dat_pen$flipper_length_mm ~ dat_pen$species)
t_test
diff_observed = round(diff(t_test$estimate), digits = 3)
print(diff_observed, digits = 3)
sum(abs(mean_differences) >= diff_observed)
```
# **Q6 (2 pts.): Given a p value of less than 1 per 10 million, how many simulations do you think you would have to do to see a difference in mean flipper length equal to or greater than 5.8 mm?**

Over 10 million, because a p-value that small indicates that we would be extremely unlikley to to see that pattern if we re-sampled due to chance alone.

# **Q7 (1 pt.): Include a boxplot of your chosen variable in your report.**
#boxplot
```{r}
boxplot(dat_pen$body_mass_g ~ dat_pen$species, xlab = "Species", ylab = "Body Mass (g)")
```

# **Q8 (3 pts.): Report the group means and difference between the means.**
#aggregate
```{r}
agg_means = aggregate(
  body_mass_g ~ species, 
  data = dat_pen, 
  FUN = mean, 
  na.rm = TRUE)
diff_observed = diff(agg_means[, 2])

agg_means
diff_crit = diff_observed
diff_crit
```
The Adelie group mean = 3700.662, the Chinstrap group mean = 3733.088. The difference between the means is 32.4259g.

# **Q9 (4 pts.): Interpret the p-value from the t-test output in plain, non-technical English that a non-statistician would understand.**

The p-value from the t-test shows that it is very likely we would observe a difference in mean body mass of 32.4259g by chance alone (i.e. because of random individual variation in our sample). If were were to re-sample this population repeatedly our random samples would have over a 50% chance of seeing this great a difference in means or greater.

```{r}
t_test = t.test(dat_pen$body_mass_g ~ dat_pen$species)
t_test
```
# **Q10 (2 pts.): How many differences in means were greater than diff_crit?**
**Remember to treat this as a 2-tailed test.**
```{r}
n = 1000
mean_differences = c()
for (i in 1:n)
{
  mean_differences = c(
    mean_differences,
    two_group_resample(dat_pen$body_mass_g, 68, 152)
  )
}
sum(abs(mean_differences) >= diff_observed)
```
I had 603 simulations where the absolute value of the difference in means were greater then or equal to the diff_crit. 


# **Q11 (2 pts.): Include a histogram of your simulation results in your report. Make sure it has appropriate title and labels.**
```{r}
hist(mean_differences, main = "Histogram of Simulated Mean Differences (Monte Carlo)", xlab = "Mean Differences")
```

