---
title: "Fassler Lab 10"
author: "Aliza Fassler"
date: "11/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
require(here)
```

On this assignment I worked with Sonja Glasser and Justin Roche.

# **Q1 (8 pts.): Submit the code you used to build your ANOVA by hand. Make sure you use the code template so that you use the same variable names as those which I’ll use for the grading.**

```{r}
rm(list = ls())

rope = read.csv(here("data", "rope.csv"))
rope$rope.type = factor(rope$rope.type)

n_obs = length(rope$rope.type)
n_groups = nlevels(rope$rope.type)

ss_tot = sum((mean(rope$p.cut) - rope$p.cut)^2)
df_tot = n_obs -1

agg_sq_resids = aggregate(
  x = rope$p.cut,
  by = list(rope$rope.type),
  FUN = function(x) sum((x - mean(x))^2))

ss_within = sum(agg_sq_resids$x)
df_within = n_obs - n_groups

ss_among = ss_tot - ss_within
df_among = df_tot-df_within


ms_among  =  ss_among / (n_groups - 1)
ms_within = ss_within / (n_obs - n_groups)

f_ratio = ms_among/ms_within
f_pval = 1 - pf(f_ratio, df_among, df_within)
```


```{r, include = FALSE}
#self check
digits_check = 5

# Build the reference model using R functions
fit_1 = lm(p.cut ~ rope.type, data=rope)
anova(fit_1)
anova_fit_1 = anova(fit_1)

# Check degrees of freedom
anova_fit_1$Df == c(df_among, df_within)

# Check sums of squares
round(anova_fit_1$`Sum Sq`, digits = digits_check) == round(c(ss_among, ss_within), digits = digits_check)

# Check mean squares
round(anova_fit_1$`Mean Sq`, digits = digits_check) == round(c(ms_among, ms_within), digits = digits_check)

# Check the F-ratio
round(anova_fit_1$`F value`[1], digits = digits_check) == round(f_ratio, digits = digits_check)

# Check the F test statistic p-value
round(anova_fit_1$`Pr(>F)`[1], digits = digits_check) == round(f_pval, digits = digits_check)
```

# **Q2 (1 pt.): Examine the conditional boxplot in the Partitioning Variance: Within-Group section of the walkthrough. Based on the figure, do you think there are equal variances among the groups?**
Based on the boxplot it looks like the rope type groups do not have equal variance. For example, the "blaze" group has a much greater spread compared to the "PI" group which has much less spread in the data and therefore less variance.

# **Q3 (1 pt.): Conduct a Bartlett test to assess the homoteneity of variances in the rope type groups. Report the p-value.**
```{r}
bartlett.test(p.cut ~ rope.type, data=rope)
```
P-value = 0.00143

# **Q4 (2 pts.): Given your graphical assessment (question 2) and the Bartlett test, do you think an ANOVA-type analysis is appropriate on the raw data? Explain why or why not.**

Based on the boxplot and significant p-value from the bartlett test, I do not think an ANOVA-type analysis is appropriate for the raw data since it violates the homogeneity assumption. Transforming the data by using for example a log transformation may stabilize the variance because the groups with higher medians seem to have more spread. 


# **Q5 (1 pt.): Which rope type is the base case?**

Blaze is the base case.

# **Q6 (1 pt.): What is the mean percent cut of the base case rope? Show your calculation using value(s) from the model coefficient table.**

The mean percent cut for the base case is the estimate = 0.36714.


# **Q7 (1 pt.): What is the mean percent cut rope type XTC? Show your calculation using value(s) from the model coefficient table.**
```{r}
 #The mean percent cut is the estimate for the base case minus the estimate for the XTC case.
0.36714-0.10164
```






