---
title: "Lab 7"
author: "Aliza Fassler"
date: "10/20/2021"
output: html_document
---
I worked with Sonja Glasser & Justin Roch on this lab.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(here)
moths = read.csv(here("data", "moths.csv"))
require(palmerpenguins)
require(boot)
```

```{r, include=FALSE}
## calculating mean bill length gentoo penguins
sse_mean = function(x)
 { 
  y = na.omit(x)
  return(sd(y)/sqrt(length(y)))
}

dat_pen = subset(penguins, species == "Gentoo")
sse_mean(dat_pen$bill_length_mm)
```


# **Q1 (1 pt.): What is the sample size, n? Show the code you used for the calculation and remember to check for missing data.**

```{r}
bill_no_na_gen<-na.omit(dat_pen$bill_length_mm)
length(bill_no_na_gen)
```

# **Q2 (1 pt.): What is the sample standard deviation? Show the code you used for the calculation.**
```{r}
sd(bill_no_na_gen)
```
# **Q3 (2 pts.): What are the critical t-values? Show the R code you used for the calculation.**
```{r}
crit_upper<-qt(0.05/2, 122, lower.tail=FALSE)
crit_lower<-qt(0.05/2, 122, lower.tail=TRUE)
crit_upper
crit_lower
```

# **Q4 (1 pt.): What is the sample standard error? Show the R code you used for the calculation.**
```{r}
sse_mean = function(x)
 { 
  y = na.omit(x)
  return(sd(y)/sqrt(length(y)))
}
gen_se <-sse_mean(bill_no_na_gen)
gen_se
```
# **Q5 (2 pts.): Finally, construct the CI and show the R code you used for the calculation.**
```{r}
upper_rad<-crit_upper*gen_se
upper_rad
lower_rad<-crit_lower*gen_se
lower_rad
mean <-mean(bill_no_na_gen)
mean+upper_rad
mean+lower_rad
```
$47.505 \pm 0.550$

# **Q6 (1 pt.): What is the CI?**
46.97236-48.06020

# **Q7 (1 pt.): Show the r code you used to call the boot() function.**
```{r}
require(boot)
```


```{r, include=FALSE}
boot_mean = function(x, i)
{
  return(mean(x[i], na.rm = TRUE))
}

```


```{r}
myboot = 
  boot(
    data = bill_no_na_gen,
    statistic = boot_mean,
    R = 10000)
print(myboot)
```
# **Q8 (12 pts.): Show the r code you used to calculate the upper and lower 2.5% quantiles.**
```{r}
quantile(
  myboot$t,
  c(0.025, 0.975))
```
# **Q9 (5 pts.): Show your completed rarefaction_sampler() function.**
```{r, include=FALSE}
moth_dat = moths[,-1]
head(moth_dat)
```



```{r}
# This clears the current R session's environment
rm(list = ls())

# Re-read my data:
moths = read.csv(here("data", "moths.csv"))
moth_dat = moths[,-1]

rarefaction_sampler = function(input_dat, n_iterations)
{
  n_input_rows = nrow(input_dat)

  results_out = matrix(
    nrow = n_iterations,
    ncol = n_input_rows)

  # The outer loop: runs once for each bootstrap iteration.  index variable is i
  for(i in 1:n_iterations)
  {
    # The inner loop: simulates increasing sampling intensity
    # Sampling intensity ranges from 1 site to the complete count of
    # sites in the input data (n)
    for(j in 1:n_input_rows)
    {
      # sample the input data row indices, with replacement
      rows_j = sample(n_input_rows, size = j, replace=TRUE)

      # Creates a new data matrix
      t1 = input_dat[rows_j, ]

      # Calculates the column sums
      t2 = apply(t1, 2, sum)

      # Counts the number of columns in which any moths were observed
      results_out[i, j] = sum(t2 > 0)
    }
  }
  return(results_out)
}

rarefact = rarefaction_sampler(moth_dat, 100)
head(rarefact)
```

# **Q10 (1 pt.): What did you find most difficult about building the function?**

The most challenging part was de-bugging and trying to see the differences between versions 1 and 2. It is hard to follow the meaning of complex functions sometimes. I ended up just trying to look at small pieces one at a time to see where it was calling something and where something was not defined within the function.

# **Q11 (4 pts.): Show the code you used to perform the simulations and construct the curve.**
```{r}
moths = read.csv(here("data", "moths.csv"))
rarefact = rarefaction_sampler(moths[,-1], 10000)
```

```{r}
rare_mean = apply(rarefact, 2, mean)
rare_quant = apply(rarefact, 2, quantile, probs=c(0.025, 0.975))
rare = t(rbind(rare_mean, rare_quant))
```

# **Q12 (4 pts.): Include your rarefaction curve plot in your report. Show the R-code you used to create your plot.**
```{r}
matplot(
  rare,
  type='l',
  col= c("blue", "purple", "red"), lwd = 6,
  xlab='Number of sampling plots',
  ylab='Species richness',
  main='Moths Rarefaction Curve')

legend(x=18, y=7.5,
  legend=c('mean','2.5%','97.5%'),
  lty=c(1,2,4),col= c("blue", "purple", "red"), lwd = 6, inset=c(.1,.1), title = "Legend", title.col = "slateblue4")
```

# **Q13 (2 pts.): About how many sites should you visit if you want to see all of the moth species? Explain your reasoning using your rarefaction curve figure.**

If you wanted to see all of the moth species you would need to visit at least ~15 plots because that is the point at which the species richness levels off. If you visit over 15 plots you are likely to see all of the species. 